---
title: "232_assignment_2"
author: "Claire Madden and Lydia Bleifuss"
date: "4/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)

clim_data <- read_csv(here("Data", "clim.csv"))
# could change this line to start from .txt
```

```{r}

# might want to consider how to do this for any climate data that might be input, in case the headers are not the same
# should just rename all the headers ?
# add error checking for column headers that are required to run the function
# check row length of columns to make sure they match 

# use sapply to say (for each year do this thing)

# pick one of the five parameters to vary - it doesn't matter which

```


1. Implement a model of almond yield anomaly (difference from average) in R based on this paper; You will be graded both on the “correctness” of the implementation and on “style”.  **Make sure you add some error checking.** Use coefficients in the paper as default values, but make sure you can vary them. 

- see .R doc

2. Have the function return almond yield anomaly for each year, and max and minimum yields over a time series of multiple year inputs

```{r}

# use function to return list of yield anomolies and the max and min anomoly over the time series

source("../R/yield_anom.R")
yield_anom_list <- yield_anom(clim_data = clim_data)

```



3. With the climate data provided (clim.txt), *use your model* to estimate how almond yield anomalies for each year in the record - graph the results.

```{r}
# first need to change the function output to a dataframe so we can graph it

yield_anom_df <- data.frame(yield_anom_list[1]) %>% 
  mutate(scenario = "baseline")


```


```{r}
yield_anom_1 <- ggplot(data = yield_anom_df, aes(x = year, y = y_anom)) +
  geom_col(fill = "peru", width = 0.5, show.legend = FALSE, alpha = 0.8) +
  theme_classic() +
   ggtitle("Yield Anom (1989 - 2010)\n") +
  xlab("\nYear") +
  ylab("Yield (tons/acre)") +
  theme(plot.title = element_text(size=12)) +
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_x_continuous(expand = c(0,0), limits = c(1989,2010)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(expand = c(0,0)) +
    labs(caption = "\nFigure 1. Caption.") 

yield_anom_1
```


4. Use model to estimate the mean annual yield anomaly if the last 20 years had been 2C warmer


```{r}

source("../R/yield_anom_2c.R")

yield_anom_list2 <- yield_anom_2c(clim_data = clim_data)

yield_anom_df2 <- data.frame(yield_anom_list2[1]) %>% 
  mutate(scenario = "2 degrees warmer") %>% 
  rename(y_anom = y_anom_2c)


mean_anom <- mean(yield_anom_df$y_anom)
mean_anom_2c <- mean(yield_anom_df2$y_anom)
```


The mean annual yield anomaly under the baseline conditions = `r mean_anom`. 
If the last 20 years had been 2 degrees warmer, the mean annual yield anomaly = `r mean_anom_2c`



```{r}

#combine yield anom for baseline and 2deg warmer into single df
yield_anom_merge <- rbind(yield_anom_df, yield_anom_df2)

# graph the results of both on single column graph
yield_anom_2 <- ggplot(data = yield_anom_merge, aes(x = year, y = y_anom))+
  geom_col(aes(fill = scenario),position = "dodge", width = 0.5, alpha = 0.8) +
  theme_classic() +
   ggtitle("Almond Yield Anomalies (1989 - 2010)\n") +
  xlab("\nYear") +
  ylab("Yield (tons/acre)") +
  theme(plot.title = element_text(size=12)) +
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_x_continuous(expand = c(0,0), limits = c(1988,2011)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(caption = "\nFigure 1. Calculated almond yield anomalies based on climate\n variables temperature and precipitation under a baseline \nscenario and under a scenario of temperature being \n2 degrees C warmer for the years 1989 - 2010.") 

yield_anom_2
```


5. Vary one of the parameter by sampling from a normal distribution with mean the value from Lobell et al., (2006) and standard deviation 10% of the mean -  create a box plot of  mean annual yield anomaly for baseline and 2C climate (where boxes show variation due to uncertanty in the parameter) 

```{r}
# trying to have a range of tn2 for each year and apply that in functions, not working for me
# 
# param_tn2 = rnorm(10, -0.0046, sd = 0.1)
# 
# 
# for(i in 1:length(param_tn2)){
# 
#   var_anom <- yield_anom(clim_data=clim_data, tn2 = param_tn2[i])
#   var_mean <- mean(data.frame((var_anom[[1]])$y_anom))
# 
#   var_anom_2c <- yield_anom_2c(clim_data=clim_data, tn2 = param_tn2[i])
#   var_2c_mean <- mean(data.frame((var_anom_2c[[1]])$y_anom_2c))
# 
# }

```


```{r}

param = rnorm(22, -0.0046, sd=0.1)
param_df = as.data.frame(param)


for(i in 1:length(param)){
  
  tn2_anom <- yield_anom(clim_data=clim_data, tn2 = param[i])
  
  tn2_anom_2c <- yield_anom_2c(clim_data=clim_data, tn2 = param[i])
  
}

```

```{r}
tn2_anom_df1 <- data.frame(tn2_anom[1]) %>% 
  mutate(scenario = "baseline")

tn2_anom_df2 <- data.frame(tn2_anom_2c[1]) %>% 
  mutate(scenario = "2 degrees warmer") %>% 
  rename(y_anom = y_anom_2c)

tn2_anom_merge <- rbind(tn2_anom_df1, tn2_anom_df2)
  
  
```


```{r}

tn2_box <- ggplot(data = tn2_anom_merge, aes(y = y_anom, x = year, fill = scenario)) +
  geom_boxplot(aes(fill = scenario)) +
  geom_point(aes(color = scenario), size = 3, alpha = 0.6) +
  theme_classic() +
  xlab("Year\n") +
  ylab("Yield Anomaly") + 
  theme(plot.title = element_text(size=10)) +
theme(plot.title = element_text(hjust = 0.5))+
 theme(legend.title = element_text(size=10), legend.position = "right")

tn2_box

```



Keep your function as a *.R file - and use an Rmarkdown document to show how you used your function to answer 3-5 

Turn both the *.R file and the Rmarkdown document on guachospace



Extra Credit - write a function that would make it easy to model ANY of the crops in Lobell 



Turn in on Gauchospace - the .R function, that you used in the project (and to create the help pages) and a text file that gives your answers to questions

