---
title: "232_assignment_2"
author: "Claire Madden and Lydia Bleifuss"
date: "4/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)

clim_data <- read_csv(here("Data", "clim.csv"))
# could change this line to start from .txt
```

```{r}

# might want to consider how to do this for any climate data that might be input, in case the headers are not the same
# should just rename all the headers ?
# add error checking for column headers that are required to run the function
# check row length of columns to make sure they match 

# use sapply to say (for each year do this thing)

# pick one of the five parameters to vary - it doesn't matter which




```


Assignment Details:

Lobell et al (2006) present several empirical models that related yield of different California crops to climate variables.

1. Implement a model of almond yield anomaly (difference from average) in R based on this paper; You will be graded both on the “correctness” of the implementation and on “style”.  **Make sure you add some error checking.** Use coefficients in the paper as default values, but make sure you can vary them. 

- see .R doc, still need to add error checking

2. Have the function return almond yield anomaly for each year, and max and minimum yields over a time series of multiple year inputs

```{r}

# use function to return list of yield anomolies and the max and min anomoly over the time series

source("../R/yield_anom.R")
yield_anom_list <- yield_anom(clim_data = clim_data)

```



3. With the climate data provided (clim.txt), *use your model* to estimate how almond yield anomalies for each year in the record - graph the results.

```{r}
# first need to change the function output to a dataframe so we can graph it

yield_anom_df <- data.frame(yield_anom_list[1])


```


```{r}
yield_anom_1 <- ggplot(data = yield_anom_df, 
       aes(x = year, y = y_anom, fill = "darksalmon")) +
  geom_col(width = 0.5, show.legend = FALSE, alpha = 0.8) +
  theme_classic() +
   ggtitle("Yield Anom (1989 - 2010)\n") +
  xlab("\nYear") +
  ylab("Yield (tons/acre)") +
  theme(plot.title = element_text(size=12)) +
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_x_continuous(expand = c(0,0), limits = c(1989,2010)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(expand = c(0,0)) +
    labs(caption = "\nFigure 1. Caption.") 

yield_anom_1
```


4. Use model to estimate the mean annual yield anomaly if the last 20 years had been 2C warmer

```{r}




```




5. Vary one of the parameter by sampling from a normal distribution with mean the value from Lobell et al., (2006) and standard deviation 10% of the mean -  create a box plot of  mean annual yield anomaly for baseline and 2C climate (where boxes show variation due to uncertanty in the parameter) 



Keep your function as a *.R file - and use an Rmarkdown document to show how you used your function to answer 3-5 

Turn both the *.R file and the Rmarkdown document on guachospace



Extra Credit - write a function that would make it easy to model ANY of the crops in Lobell 



Turn in on Gauchospace - the .R function, that you used in the project (and to create the help pages) and a text file that gives your answers to questions



Useful R command -group_by


**Hint**

clim_month = clim %>% group_by(month, year) %>% summarize(meantmax = mean(tmax_c), precip=sum(precip)) 

gives average monthly maximum daily temperature, and monthly precipitation from a data frame called clim  with columns year, month, precip and tmax_c


