---
title: "sample_code_functions"
author: "Lydia Bleifuss"
date: "4/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(knitr)
```

Working with parameters 

```{r solar}
source("../R/solarpv.R")
solarpv

# read in R formatted data
load("../Data/sierraczosolar.rda")

# already in the format required for the model
head(sierraczosolar)

# run the model
solarpv(area=0.1, solar=sierraczosolar, clr="green", eunit="W")

# run and save results - but don't plot
site1 = solarpv(area=0.1, solar=sierraczosolar, clr="green", eunit="W", g=FALSE)
site1$mean

# consider a different pv array that only uses 
# diffuse and has non standard efficiency (0.6)
site2 = solarpv(area=0.1, solar=sierraczosolar, clr="green", eunit="W", g=FALSE, eff=0.6, etype="direct")
site2$mean

# try some additional variations
# add an output that is the minimum annual solar output
# add error checking that would remove years that have less than 12 months of data

```

Nested Functions + Graphing

```{r valuep}

source("../R/compute_NPV.R")
source("../R/compute_profit_frompower.R")

compute_profit_frompower

compute_NPV

# use solar from  site1 from above as input

profit_solar = compute_profit_frompower(energy=site1$annual$elect, price=30, discount=0.04)
profit_solar

ggplot(profit_solar$yearprofit, aes(year, netpre))+geom_col()+labs(y="Profit in current $", x="year")

# notice that by graphing we identify an issue with the first year
# thoughts on what it might be
```


Creating a Sample 

```{r reuse}

# rerun 'sampling' in makingfunction1.Rmd. above to get results from our reserovir model
source("../R/power_gen.R")
number_runs = length(profit_solar$year)
reservoir_model_res= as.data.frame(matrix(nrow=number_runs, ncol=3))
colnames(reservoir_model_res)=c("height","flowrate","power")
# STep 2 - generate heights and flow rates
reservoir_model_res$height = rnorm(mean=10, sd=1, n=number_runs)
reservoir_model_res$flowrate = runif(min=0.1, max=1, n=number_runs)
# Step 3 - apply model to get power for each height, flow rate
reservoir_model_res$power = mapply(FUN=power_gen, height=reservoir_model_res$height, flow=reservoir_model_res$flowrate)
head(reservoir_model_res)

# create profit from hydro
profit_hydro = compute_profit_frompower(energy=reservoir_model_res$power, price=30, discount=0.04)

# compare with
profit_solar = compute_profit_frompower(energy=site1$annual$elect, price=30, discount=0.04)

profit_solar$total/1000
profit_hydro$total/1000

```
MORE SAMPLING IN MAKING FUNCTIONS 2



Catagorical Data

R uses **factors** to deal with categorical data
and **summary** to get counts (occurence of unique entities)


# Lists - For returning more than one item from your function

*Lists are the most “informal” data structures in R
*List are really useful for keeping track of and organizing groups of things that are not all the same
*A list could be a table where number of rows is different for each column
*A list can have numeric, character, factors all mixed together
* List are often used for returning more complex information from function (e.g. lm)


Example: Lets build on our compute_simpson_index, but output some additional information - the most frequent species and the rarest species

useful R
recall for *factors* , *summary* gives you occurance for each unique factor

we will also look at some other useful commands for lists

* *lapply

```{r usinglists}




# use of summary with factos
flower_types = c("iris","daisy","poppy","rose","dandelion","weed","violet")
# sample (with replacement)
garden = sample(flower_types, size=20, replace=T)

# get frequencies
summary(as.factor(garden))

# find most frequent factor
which.max(summary(as.factor(garden)))

# get the label/name of least frequent factor
names(which.min(summary(as.factor(garden))))


# use these to inclue in our function and return more # information

source("../R/compute_diversity.R")

compute_diversity

compute_simpson_index(garden)
compute_diversity(garden)

# modify to also output least frequent species and total number of unique speceis and add some error checking!

# a bit more on using functions for multiple values
# generate data for multiple gardens in multiple sites

# with lists can mix arrays and single values
site = c("A","B","C","D","E")

# generate sample data
nflowers=20
garden_sample = sapply(site, function(x) {flowers = sample(flower_types, size=nflowers, replace=T); return(flowers)})

# or with a for loop
garden_sample = matrix(nrow=nflowers, ncol=length(site))
for (j in 1:length(site)) {
  garden_sample[,j] = sample(flower_types, size=nflowers, replace=T)
}

# example application of our function to a specific garden
compute_diversity(garden_sample[,2])

# now generate a data frame to store our results for all gardens
# recall for each garden we now have multiple things returned
# a) the simpson index and b) the most frequent species


many_garden_result = data.frame(site = c("A","B","C","D","E"), simpson=NA, mostfreq=NA)

many_garden_result
for (i in 1:length(site)) {
  tmp = compute_diversity(garden_sample[,i])
  many_garden_result$simpson[i]=tmp$simpson
  many_garden_result$mostfreq[i]=tmp$mostfreq
}
many_garden_result

# alteratively - a bit cleaner but runs function twice
for (i in 1:length(site)) {
   many_garden_result$simpson[i]=compute_diversity(garden_sample[,i])$simpson
  many_garden_result$mostfreq[i]=compute_diversity(garden_sample[,i])$mostfreq
}

head(many_garden_result)
ggplot(many_garden_result, aes(site, simpson, fill=mostfreq))+geom_col()

# we could also use apply to run our model over all of our sites
# apply lets you run for each column
# it will result in list
# advantage is you don't need to define the data structure
# disadvantage lists are hard to deal with

many_gardens_result_aslist = apply(garden_sample, 2, function(x) {diversity= compute_diversity(x); return(diversity)})

head(many_gardens_result_aslist)

# look at an individual garden (element of list)
many_gardens_result_aslist[3]
# all named items
lapply(many_gardens_result_aslist,"[", c("simpson"))

# often easier to get it out of list format
# organize list by columns and convert to dataframe
many_gardens_result_aslist[1]
# note that each element of the list has 2 items- so store as a matrix with 2 columns one for each item - and use unlist to extract

many_gardens_sum = as.data.frame(matrix(unlist(many_gardens_result_aslist), ncol=2, byrow=T))
colnames(many_gardens_sum)=c("simpson","mostfreq")
# add site name
many_gardens_sum$site = site


#plot
ggplot(many_gardens_sum, aes(site, simpson, fill=mostfreq))+geom_col()+labs(y="Simpson Index", x="Site")

#try adding least frequent species and total number of species to the function

```


